### 2023-01-23 article3.qmd
update article 3

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux. During these drift phases, measurements of particles abundance (between 0.6 mm - 2.5 mm) are taken every 30 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication.

```{r, init function}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
```

```{r, map of 4 floats}
#| label: fig-float-position
#| fig-cap: Route of each float since their deployment (black triangles). Dots represent ascending profiles. The profile that concludes the time series for each float is highlighted by a black contour.
#library(tidyverse)

@@

We applied a similar approach to @Estapa2013-gw and @Estapa2017-jr to derive the continuous component (i.e. continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e. deposition of stochastic big particles) components of the downward POC flux.

First, the drifting $c_{p}$ signal was cleaned by removing spikes (hypothesized as transient particles not landing on the window or active swimmers [@Estapa2013-gw]) using the despike function (oce package in R, see https://dankelley.github.io/oce) where a spike is defined here as any value higher than 4 times the standard deviation of the difference between raw $c_{p}$ data and a smoothed (i.e. reference) $c_{p}$ signal, obtained with a 7-point running median filter. Despiked $c_{p}$ data were then smoothed with a 3-point running median filter to ease the detection of jumps, defined as a continuous increase or decrease of $c_{p}$ (@fig-estapa-method_article).

```{r, fig fit}
#| label: fig-estapa-method_article
#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component and a discontinuous (positive jumps, red) component. Slopes for each group are depicted in light blue. A negative jump was also detected (yellow). Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.
clean_cp_data <- function(data, moving_median_order){

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach to detect jumps. For this, we applied a 7-point running slope on drifting $c_{p}$ data at each drifting depth and we then computed the 90$^{th}$ quantile ($Q_{90}$) of each slope distribution with a 10-day increasing dataset since the deployment (Fig-XX) to determine the most appropriate slope thresholds to detect jumps. The latter are however subject to a tradeoff between the likely underestimation of $F_{small}$ and the detection of false jumps (without great impact on the value of $F_{big}$ because false jumps have low $c_{p}$ values) if the threshold is too small and the likely overestimation of $F_{small}$ to the detriment of $F_{big}$ with missed jumps if the threshold is too high. As shown in Fig-XX, the value of $Q_{90}$ decreases over time for each drifting depth. This is most likely explained by the vanishing of big particles (> 0.5 mm in this study) over time (Fig-YY). Therefore, we assumed that past a given time, the great majority of jumps have been seen by the transmissometer and that the optimal values for each slope threshold should be found in the period following the bloom up to this given time. We thus have found that on a weekly basis, particles in the 0.5 - 1.5 mm range at 1000 m were starting to steadily decrease (assumed to be slowly removed without upwards production) around week 40, that is the beginning of October. (maybe find another criteria .. thresholds as they currently are give good fit.)

```{r, slope distribution}

# remove bad data from float 6904241
float_6904241_slope <- float_6904241 |> filter(cycle < 20)

# create dataframe for all drifting data (for all floats)
all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)

# compute slope for every float, with every size window
wmo <- c(4903634, 6904240, 6904241, 1902578)
park_depth <- c('200 m', '500 m', '1000 m')
#window_size <- c(3,5,7)
window_size <- 7

# function to clean cp data
clean_cp_data <- function(data, moving_median_order){
  if(nrow(data) == 0){ # no cp data
    return(data)
  }else{
    # despike cp data
    data$cp <- oce::despike(data$cp, reference = 'median', k = 7, replace = 'NA')
    # median average
    data$cp <- pastecs::decmedian(data$cp, order = moving_median_order, ends = "fill") |> pastecs::extract(component = "filtered")
  }
  return(data)
}

# compute cp slope
delta_y <- function(x, window_size){
  delta_y <- x[window_size]-x[1]
  return(delta_y)
}

delta_x <- function(x, window_size){
  delta_x <- (x[window_size]-x[1])/86400 # units in days
  return(delta_x)
}

compute_slope <- function(data, window_size){
  tmp <- data |> drop_na(cp)
  despiked_data <- clean_cp_data(tmp, 3) 
  if(window_size == 3){
    despiked_data$slope <- slide(despiked_data$cp, 1, delta_y, window_size = 3)/slide(despiked_data$dates, 1, delta_x, window_size = 3)
  }else if(window_size == 5){
    despiked_data$slope <- slide(despiked_data$cp, 2, delta_y, window_size = 5)/slide(despiked_data$dates, 2, delta_x, window_size = 5)
  }else{
    despiked_data$slope <- slide(despiked_data$cp, 3, delta_y, window_size = 7)/slide(despiked_data$dates, 3, delta_x, window_size = 7)
  }
  return(despiked_data)
}

res <- data.frame()
for (i in wmo){
  #print(i)
  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
  for (j in park_depth){
    for (k in seq(1:max_cycle)){
      #print(k)
      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
      if(nrow(tmp) == 0){
        next
      }else if(nrow(tmp) < 3){ # case where there is not enough data
        next
      }else{
        cleaned_signal <- clean_cp_data(tmp, 3)
        if(nrow(cleaned_signal) == 0){
          next
        }else{
          for (l in window_size){
            test <- compute_slope(cleaned_signal, l)
            test <- test |> select(dates, wmo, cycle, park_depth, slope)
            # add time_drifting
            max_time <- max(test$dates)
            min_time <- min(test$dates)
            drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
            test$drifting_time <- drifting_time
            test$window_size <- l
            res <- rbind(res, test)
          } 
        }
      }
    }
  }
}

fixed_dates <- c('2022-06-01', '2022-06-10', '2022-06-20', 
                 '2022-07-01', '2022-07-10', '2022-07-20', 
                 '2022-08-01', '2022-08-10', '2022-08-20', 
                 '2022-09-01', '2022-09-10', '2022-09-20', 
                 '2022-10-01', '2022-10-10', '2022-10-20', 
                 '2022-11-01', '2022-11-10', '2022-11-20', 
                 '2022-12-01', '2022-12-10', '2022-12-20',
                 '2023-01-01')

res2 <- data.frame()
for (i in fixed_dates){
  tmp <- filter(res, dates < i)
  tmp2 <- tmp |> group_by(park_depth, window_size) |> dplyr::summarise(threshold = quantile(slope, 0.90, na.rm = T), nb_points = n())
  tmp2$group <- i
  tmp2$quantile <- 90
  res2 <- rbind(tmp2, res2)
}

# add levels
res2$park_depth <- factor(res2$park_depth, levels = c('200 m', '500 m', '1000 m'))

# plot result
# res3 <- res2 |> mutate(juld = as_date(group)) |> ungroup()
# ggplot(res3, aes(x = juld, y = threshold, colour = factor(window_size), group = factor(window_size))) + geom_point() + geom_path() + 
#   facet_wrap(~park_depth, scales = 'free') + theme_bw() +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   scale_color_brewer(palette = 'Dark2') +
#   labs(x = 'Month', y = 'Slope')
res3 <- res2 |> mutate(juld = as_date(group)) |> ungroup()
ggplot(res3, aes(x = juld, y = threshold)) + geom_point() + geom_path() + 
  facet_wrap(~park_depth, scales = 'free') + theme_bw() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  geom_vline(xintercept = as_date('2022-10-01'), colour = "red", linetype = 'dashed') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  scale_color_brewer(palette = 'Dark2') +
  labs(x = 'Month', y = 'Slope')

As a result, we have chosen to set the slope thresholds to detect jumps at week 40 for all three parking depths at, in increasing order with depth, 0.23, 0.12 and 0.10. Using those thresholds, we computed the slope between each $c_{p}$ measurement of each drifting periods and identified as jump any point above the selected thresholds. This method allowed us to separate the continuous component of the downward POC flux into groups, between each jump. We then computed the slope for each group (light blue straight lines in @fig-estapa-method_article). Any negative slope was disregarded assuming that either the so-called fit failed (ineffective threshold leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window. Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each group in order to derive the small particles flux ($F_{small}$) using the relationship of @Estapa2022

$$
F_{small} = 633(0.25~slope_{w})^{0.77}
$$ {#eq-estapa}

The discontinuous component of the downward POC flux or big particle flux ($F_{big}$) was derived using @eq-estapa by replacing $slope_{w}$ as the sum of positive jumps (negative jumps were disgarded) divided by the drifting time.

All profiles analyzed with this method are available at https://fricour.shinyapps.io/CARBON_REFINE/.

@@

Using the selected slope thresholds determined in the previous section, we computed $F_{small}$ and $F_{big}$ at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between 5.1 and 26.9 mg C m$^{-2}$ day$^{-1}$ with an average of 14.1 $\pm$ 3.9 mg C m$^{-2}$ day$^{-1}$ (standard deviation). At 500 m, $F_{small}$ is comprised between 6.9 and 19.8 mg C m$^{-2}$ day$^{-1}$ with an average of 9.5 $\pm$ 2.1 mg C m$^{2}$ day$^{-1}$. At this depth, we observe a global decrease of $F_{small}$ with time though float 6904240 measured a relatively constant downward $F_{small}$ at 8.6 mg C m$^{-2}$ day$^{-1}$ on average. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards December, with $F_{small}$ ranging between 4.0 and 11.8 mg C m$^{-2}$ day$^{-1}$ and 7.0 $\pm$ 1.4 mg C m$^{-2}$ day$^{-1}$ on average.

```{r, function plot jump}
# plot_jumps <- function(clean_data, threshold){
#   
#   # save bbp data for later
#   #bbp_df <- clean_data |> select(dates, bbp) |> drop_na(bbp)
#   
#   # save cleaned Cp signal from spikes to a temporary variable
#   tmp2 <- clean_data |> arrange(dates)
#   #print(colnames(tmp2))
#   # init jump values to NA
#   tmp2$jump <- NA
#   # compute slope between two adjacent points (except first and last point of drifting time series)
#   for (i in 2:(nrow(tmp2)-1)){
#     #print(i)
#     delta_x <- as.numeric(difftime(tmp2$dates[i], tmp2$dates[i-1], units = 'days'))
#     delta_y <- tmp2$cp[i]-tmp2$cp[i-1]
#     tmp2$jump[i] <- delta_y/delta_x
#   }
#   
#   # assign a colour for the plot
#   tmp2$colour <- 'base signal'
#   tmp2$colour[which(abs(tmp2$jump)>threshold)] <- 'jump' # abs is needed for 'negative jumps'
#   
#   # add group to compute slope (for each subgroups of points, separated from a jump)
#   tmp2$group <- NA
#   
#   # compute where the jumps are (index value of the array)
#   jump_index <- which(tmp2$colour == 'jump')
#   
#   # assign group 
#   for (i in jump_index){
#     for (j in 1:nrow(tmp2)){
#       if ((j < i) & (is.na(tmp2$group[j]))){
#         tmp2$group[j] <- paste0('group_',i)
#       }
#     }
#   }
#   
#   # add last group (because it is AFTER the last index so it was not computed before)
#   tmp2$group[which(is.na(tmp2$group))] <- 'last_group'
#   
#   # compute slope for each subgroups
#   slope_df <- tmp2 |> filter(colour == 'base signal', jump != 'NA') |> dplyr::group_by(group) |> dplyr::summarise(min_time = min(dates), max_time = max(dates), 
#                                                                                                                   nb_points = n(), first_cp = cp[1], last_cp = cp[nb_points],
#                                                                                                                   delta_x = as.numeric(difftime(max_time, min_time, units = 'days')),
#                                                                                                                   delta_y = (last_cp-first_cp)*0.25, slope = delta_y/delta_x)
#   
#   # remove negative slope from the mean slope (no physical meaning)
#   slope_df <- slope_df |> filter(slope > 0)
#   
#   # remove if only one point (cannot fit a slope with one point)
#   slope_df <- slope_df |> filter(nb_points > 1)
#   
#   # compute mean slope 
#   #mean_slope <- mean(slope_df$slope)
#   mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)
#   
#   # estapa relationship (see POSTER)
#   #poc_flux <- 633*((mean_slope*0.25)**0.77) # *0.25 because ATN = cp*0.25
#   poc_flux <- 633*((mean_slope)**0.77) # /!\ slope for computed for ATN on y axis (delta_y *0.25 because ATN  = cp*0.25) -> should be OK
#   
#   # build dataframe to plot each subgroup
#   part1 <- slope_df |> select(group, time = min_time, cp = first_cp)
#   part2 <- slope_df |> select(group, time = max_time, cp = last_cp)
#   part_slope <- rbind(part1, part2)
#   
#   # spot negative jump
#   tmp2$colour[which((tmp2$colour == 'jump') & (tmp2$jump < 0))]  <- 'negative jump'
#   
#   # add big particles flux
#   rows_to_keep <- c(jump_index, jump_index-1) 
#   tmp3 <- tmp2[rows_to_keep,] |> select(dates, cp, jump, colour, group) |> dplyr::arrange(dates)
#   
#   # remove negative jump, if any
#   check_colour <- unique(tmp3$colour)
#   if(length(check_colour) == 3){
#     index_neg_jump <- which(tmp3$colour == 'negative jump')
#     tmp3 <- tmp3[-c(index_neg_jump, index_neg_jump+1),]
#     tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
#     even_indexes <- seq(2,nrow(tmp3),2)
#     tmp3 <- tmp3[even_indexes,]
#   }else if(length(check_colour) == 2){
#     tmp3 <- tmp3 |> mutate(diff_jump = cp - lag(cp)) 
#     even_indexes <- seq(2,nrow(tmp3),2)
#     tmp3 <- tmp3[even_indexes,]
#   }else{ # NO jump
#     tmp3 <- NULL
#   }
#   
#   # big particles flux
#   if(is.null(tmp3)){ # no jumps
#     big_part_poc_flux <- 0
#   }else{
#     tmp4 <- tmp3 |> filter(diff_jump > 0)
#     if(nrow(tmp4) == 0){ # no positive jumps
#       big_part_poc_flux <- 0
#     }else{
#       delta_y <- sum(tmp4$diff_jump) *0.25 # to get ATN (= cp*0.25)
#       max_time <- max(tmp2$dates)
#       min_time <- min(tmp2$dates)
#       delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
#       slope_big_part <- delta_y/delta_x
#       big_part_poc_flux <- 633*(slope_big_part**0.77)
#     }
#   }
#   
#   # compute total drifting time
#   max_time <- max(tmp2$dates)
#   min_time <- min(tmp2$dates)
#   drifting_time <- as.numeric(difftime(max_time, min_time, units = 'days'))
#   
#   jump_plot <- ggplot(clean_data, aes(x = dates, y = cp)) + geom_point(size= 3) + geom_point(data = tmp2, aes(x = dates, y = cp, colour = colour), size = 3) + 
#     scale_colour_manual(values = c('#003366','#E31B23', '#FFC325')) + 
#     scale_shape_manual(values = c(1,1,4)) +
#     geom_path(data = part_slope, aes(x = time, y = cp, group = group), colour = '#DCEEF3', size = 2) + theme_bw() + 
#     labs(x = 'Time', y = 'Cp (1/m)', title = paste0('Drifting time: ', round(drifting_time,3), ' days\n',
#                                                     'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n', 
#                                                     'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
#                                                     'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1')) #+
#   #geom_point(data = bbp_df, aes(x = dates, y = bbp), colour = "pink")
#   
#   
#   # adapt script to return big and small flux
#   df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = clean_data$park_depth[1], wmo = clean_data$wmo[1],
#                cycle = clean_data$cycle[1])
#   
#   #return(jump_plot)
#   #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))
#   return(df)
#   
# }

@@

  # despike cp data with a 7-point moving window
  tmp$cp <- despike(tmp$cp, k = 7)

@@

  # spot outliers on the Z score signal
  # interquartile range between Q25 and Q75
  IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
  # outliers ('spikes' in the Z score signal)
  spikes_down <- tmp$zscore < quantile(tmp$zscore, 0.25, na.rm=T) - 1.5 *IQR
  spikes_up <-  tmp$zscore > quantile(tmp$zscore, 0.75, na.rm=T) + 1.5 *IQR
  spikes <- as.logical(spikes_down + spikes_up)

@@

  # compute weighted average slope (to take into account the fact that some subgroup might have 2 points and a high slope vs. large group of points with a small slope)
  mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)

  # convert cp to POC using Estapa's relationship 
  poc_flux <- 633*((mean_slope)**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK

@@

  # remove negative jump, if any
  check_colour <- unique(tmp2$colour)
  if(length(check_colour) == 3){ # there is a negative jump
    index_neg_jump <- which(tmp2$colour == 'negative jump')
    tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
    even_indexes <- seq(2,nrow(tmp2),2)
    tmp2 <- tmp2[even_indexes,]
  }else if(length(check_colour) == 2){ # only positive jump
    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
    even_indexes <- seq(2,nrow(tmp2),2)
    tmp2 <- tmp2[even_indexes,]
  }else{ # No jump
    tmp2 <- NULL
  }

  if(is.null(tmp2)){ # no jumps
    big_part_poc_flux <- 0
  }else{
    tmp3 <- tmp2 |> filter(diff_jump > 0)
    if(nrow(tmp3) == 0){ # no positive jumps
      big_part_poc_flux <- 0
    }else{
      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
      max_time <- max(tmp$dates)
      min_time <- min(tmp$dates)
      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
      slope_big_part <- delta_y/delta_x
      big_part_poc_flux <- 633*(slope_big_part**0.77)
    }
  }

@@

  jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
    add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
    layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
                         'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
                         'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
                         'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))

@@

    # adapt script to return big and small flux
  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = clean_data$park_depth[1], wmo = clean_data$wmo[1],
               cycle = clean_data$cycle[1])

  #return(jump_plot)
  #return(list('jump_plot' = jump_plot, 'jump_table' = tmp3))
  return(df)

@@

```{r}
#| label: fig-OST-flux
#| fig-cap: OST-derived small (top) and big (bottom) particles flux for all floats at each drifting depth. Black lines and shaded areas represent, respectively, the monthly mean flux and the 95% confidence interval around the mean.

@@

res4 <- data.frame()
for (i in wmo){
  #print(i)
  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
  for (j in park_depth){
    for (k in seq(1:max_cycle)){
      #print(k)
      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
      if(nrow(tmp) == 0){
        next
      }else if(nrow(tmp) < 3){ # case where there is not enough data
        next
      }else{
        cleaned_signal <- clean_cp_data(tmp, 3)
        if(nrow(cleaned_signal) == 0){
          next
        }else{
          #for (l in window_size){
            #print(l)
            #for (m in quartile){
              #print(m)
          if(j == '200 m'){
            slope_threshold <- 0.23
          }else if(j == '500 m'){
            slope_threshold <- 0.12
          }else{
            slope_threshold <- 0.1
          }
          #selected_threshold <- threshold_table |> filter(park_depth == j, window_size == l, quartile == m)
          #slope_threshold <- selected_threshold$slope_threshold
          output <- plot_jumps(cleaned_signal, slope_threshold)
          #print(output)
          # append result
          res4 <- rbind(res4, output)
            #}
          #} 
        }
      }
    }
  }
}

# keep the good fluxes
cflux <- res4 |> mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))

info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
                                                                mean_smallf = mean(small_flux, na.rm=T), median_smallf = median(small_flux, na.rm=T),
                                                                std_smallf = sd(small_flux, na.rm=T),
                                                                min_bigf = min(big_flux, na.rm=T), max_bigf = max(big_flux, na.rm=T),
                                                                mean_bigf = mean(big_flux, na.rm=T), median_bigf = median(big_flux, na.rm=T),
                                                                std_bigf = sd(big_flux, na.rm=T))

@@

###################################
# MOST RECENT VERSION OF THE PLOT #
###################################

up <- cflux |> ggplot(aes(x = date, y = small_flux)) +
  geom_smooth(data = cflux, aes(x = date, y = small_flux), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + #facet_grid(~park_depth)
  #scale_x_date(labels = date_format("%b"), date_breaks = '1 month') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  #theme(axis.text.x = element_text(angle = 45, vjust = 0, hjust=0)) +
  guides(fill=guide_legend(title="WMO")) +
  #geom_pointrange(data = info_table2, aes(x = month, y = mean_small_f, ymin = mean_small_f - sd_small_f, ymax = mean_small_f + sd_small_f))
  #theme_bw() +
  geom_point(data = cflux, aes(x = date, y = small_flux, colour = WMO, shape = WMO)) +
  facet_wrap(~park_depth, scales = 'free') +
  labs(x = '', y = TeX('$F_{small}$ (mg C m$^{-2}$ day$^{-1}$)')) +
  theme_bw() #+
  # theme(axis.ticks.x = element_blank(),
  # axis.text.x = element_blank(), legend.position = 'top')
  #theme(legend.position = '')

>>>>

All 4 floats have been configured to drift at 3 distinct parking depths (respectively, 200 m, 500 m and 1000 m for approximately 1, 3 and 5 days) to study the downward carbon flux. During these drift phases, measurements of particles abundance (14 size classes, between 0.1 mm and 2.5 mm) are taken every 20 minutes at 200 m and every 2h at 500 m and 1000 m whereas $c_{p}$ measurements are taken every 30 minutes at all depths. This configuration can be adjusted during the mission of the float via two-way Iridium communication.

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
```

```{r, map of 4 floats}
#| label: fig-float-position
#| fig-cap: Route of each float since their deployment (black triangles). Dots represent ascending profiles. The profile that concludes the time series for each float is highlighted by a black contour.

@@

We applied a similar approach to @Estapa2013-gw and @Estapa2017-jr to derive the continuous component (i.e. continuous deposition of small particles on the upward-facing window of the transmissometer) and discontinuous (i.e. deposition of stochastic big particles) components of the downward POC flux, hereafter referred to as, respectively, $F_{small}$ and $F_{big}$.

First, the drifting $c_{p}$ signal was cleaned from spikes (hypothesized as transient particles not landing on the window or active swimmers [@Estapa2013-gw]). For this, we computed a 7-point moving median and a 7-point moving median absolute deviation (MAD). Then, we computed the anomaly to the moving median as the difference between the drifting $c_{p}$ signal and the moving median. We considered that a $c_{p}$ measurement was a spike when the anomaly was above $5.2 \times MAD$ following @Leys2013. Despiked $c_{p}$ data were then smoothed with a 7-point moving median filter.

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. As a result, we firstly computed the slope between each point, then we computed the z-score (i.e. standard score) of each slope assuming a normal distribution of the latter. The detection of jumps thus consisted in spotting outliers in the z-score distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

In order to compute $F_{small}$, we firstly computed the slope of each group of points between jumps (positive and negative, see @fig-estapa-method_article). Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window. Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the relationship of @Estapa2022

$$
F_{small} = 633(0.25~slope_{w})^{0.77}
$$

We used the same relationship to compute $F_{big}$ was replacing $slope_{w}$ as the sum of positive jumps (negative jumps were disgarded) divided by the drifting time.

```{r, example of a cp signal analyzed for jumps (with an old method here)}
#| label: fig-estapa-method_article
#| fig-cap: Application of the method of @Estapa2017-jr on a drifting transmissometer to divide the $c_{p}$ signal into a continuous (dark blue) component and a discontinuous (positive jumps, red) component. Slopes for each group are depicted in light blue. A negative jump was also detected (yellow). Data from BGC-ARGO float WMO 6904240, drifting at 1000 m during its 21$^{st}$ cycle.
clean_cp_data <- function(data, moving_median_order){

@@

@@

```{r, function to detect and plot jumps}

@@

  # despike cp data with a 7-point moving window
  tmp$cp <- despike(tmp$cp, k = 3)

@@

  # spot outliers on the Z score signal
  # interquartile range between Q25 and Q75 -> had to used that and not the despike function because slopes are often close (or equal) to 0 so it can miss clear jumps. Q25 and Q75 are more trustworthy in this case than the despike function of Jean-Olivier (see package castr on his github: https://github.com/jiho/castr)
  IQR <- quantile(tmp$zscore, probs = 0.75, na.rm=T) - quantile(tmp$zscore, probs = 0.25, na.rm=T)
  # outliers ('spikes' in the Z score signal)
  spikes_down <- tmp$zscore < quantile(tmp$zscore, 0.25, na.rm=T) - 1.5 *IQR
  spikes_up <-  tmp$zscore > quantile(tmp$zscore, 0.75, na.rm=T) + 1.5 *IQR
  spikes <- as.logical(spikes_down + spikes_up)

@@

  # compute weighted average slope (to take into account the fact that some subgroups might have 2 points and a high slope vs. large group of points with a small slope)
  mean_slope <- sum(slope_df$nb_points * slope_df$slope)/sum(slope_df$nb_points)

  # convert cp to POC using Estapa's relationship 
  poc_flux <- 633*(mean_slope**0.77) # /!\ slope computed for ATN on y axis (delta_y *0.25 because ATN = cp*0.25) -> should be OK

@@

  # remove negative jumps, if any
  check_colour <- unique(tmp2$colour)
  if(length(check_colour) == 3){ # there is at least a negative jump
    index_neg_jump <- which(tmp2$colour == 'negative jump')
    tmp2 <- tmp2[-c(index_neg_jump, index_neg_jump+1),]
    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
    even_indexes <- seq(2,nrow(tmp2),2)
    tmp2 <- tmp2[even_indexes,]
  }else if(length(check_colour) == 2){ # only positive jumps
    tmp2 <- tmp2 |> mutate(diff_jump = cp - lag(cp)) 
    even_indexes <- seq(2,nrow(tmp2),2)
    tmp2 <- tmp2[even_indexes,]
  }else{ # No jump
    tmp2 <- NULL
  }

  if(is.null(tmp2)){ # no jump
    big_part_poc_flux <- 0
    tmp3 <- NULL
  }else{
    tmp3 <- tmp2 |> filter(diff_jump > 0)
    if(nrow(tmp3) == 0){ # no positive jumps
      big_part_poc_flux <- 0
    }else{
      delta_y <- sum(tmp3$diff_jump) *0.25 # to get ATN (= cp*0.25)
      max_time <- max(tmp$dates)
      min_time <- min(tmp$dates)
      delta_x <- as.numeric(difftime(max_time, min_time, units = 'days'))
      slope_big_part <- delta_y/delta_x
      big_part_poc_flux <- 633*(slope_big_part**0.77)
    }
  }

@@

  # plot
  jump_plot <- plot_ly(tmp, x = ~dates, y = ~cp, type = 'scatter', mode = 'markers', color = ~colour, colors = c('#003366','#E31B23', '#FFC325')) |>
    add_lines(data= part_slope_tmp, x = ~dates, y = ~cp, split = ~group, color = I('#DCEEF3'), showlegend = F) |>
    layout(title= paste0('Drifting time: ', round(drifting_time,3), ' days\n',
                         'Mean ATN slope (light blue): ', round(mean_slope,3), ' day-1\n',
                         'POC flux (small particles): ', round(poc_flux,1), ' mg C m-2 day-1\n',
                         'POC flux (big particles): ', round(big_part_poc_flux,1), ' mg C m-2 day-1'), yaxis = list(title = 'Cp (1/m)'), xaxis = list(title = 'Time'))

@@

    # adapt script to return big and small flux
  df <- tibble('max_time' = max_time, 'min_time' = min_time, 'small_flux' = poc_flux, 'big_flux' = big_part_poc_flux, park_depth = data$park_depth[1], wmo = data$wmo[1],
               cycle = data$cycle[1])

  return(df)

@@

```{r, compute Fsmall and F big}

# remove bad data from float 6904241
float_6904241_slope <- float_6904241 |> filter(cycle < 20)
# create dataframe for all drifting data (for all floats)
all_floats_slope <- rbind(float_4903634, float_6904240, float_6904241_slope, float_1902578)
all_floats_slope <- all_floats_slope |> filter(PhaseName == 'PAR') |> drop_na(cp) |> select(depth = pres, cp, dates = juld, park_depth, wmo, cycle)

@@

res <- data.frame()
for (i in wmo){
  max_cycle <- as.numeric(all_floats_slope |> filter(wmo == i) |> dplyr::summarise(max_cycle = max(cycle)))
  for (j in park_depth){
    for (k in seq(1:max_cycle)){
      tmp <- all_floats_slope |> filter(wmo == i, park_depth == j, cycle == k)
      if(nrow(tmp) == 0){
        next
      }else if(nrow(tmp) < 3){ # case where there is not enough data
        next
      }else{
          output <- plot_jumps(tmp)
          res <- rbind(res, output)
      }
    }
  }
}

# keep the good fluxes
cflux <- res |> mutate(drifting_time = difftime(max_time, min_time, units = 'days'), WMO = factor(wmo))

cflux_info_table <- cflux |> dplyr::group_by(park_depth) |> summarize(min_smallf = min(small_flux, na.rm=T), max_smallf = max(small_flux, na.rm=T),
                                                                mean_smallf = mean(small_flux, na.rm=T), median_smallf = median(small_flux, na.rm=T),
                                                                std_smallf = sd(small_flux, na.rm=T),
                                                                min_bigf = min(big_flux, na.rm=T), max_bigf = max(big_flux, na.rm=T),
                                                                mean_bigf = mean(big_flux, na.rm=T), median_bigf = median(big_flux, na.rm=T),
                                                                std_bigf = sd(big_flux, na.rm=T))

@@
